---
title: "Homework 3"
author: "Henry Stoddard"
date: "9/24/2020"
output: github_document
---
## Problem 0
Creating initial markdown file and rendering to github doc

## Problem 1
```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns.
Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables - name, asile, department, and some numeric codes.

How many aisles and which are most items from?

```{r}
instacart %>% count(aisle) %>% arrange(desc(n))
```

Making a plot!
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table!!

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

